Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-15 Ben_Manuscripts/structure_paper/Draft.tex
New figures and plotting capabilities
polishing text

  \section*{Introduction}

  Nanostructured membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control membrane architecture at the atomic scale
  permitting the design of solute-specific separation membranes.
  \cite{humplik_nanostructured_2011} Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF). \cite{van_der_bruggen_review_2003}

  While RO and NF have seen many advances in the past few decades, they are far
  from perfect separation technologies. Current commericial RO membranes are thin
  film composite membranes with a porous support layer and an active layer made
  of a dense polyamide polymer matrix.  The membranes are unstructured with
  tortuous and polydisperse diffusion pathways. Separation occurs as solutes
  dissolve into and diffuse through voids in the polymer matrix at different
  rates.\cite{wijmans_solution-diffusion_1995} With optimization, one can exploit
  these differences to create a functional selective barrier. High feed pressures
  are required in order to achieve economical solvent flux. The largest
  application of reverse osmosis is desalination. A typical seawater desalination
  process requires feed pressures in the range of 800 to 1000 psi in order to
  overcome osmotic pressure. \cite{fritzmann_state---art_2007} Only 0.5 \% of the
  world's water is fresh. Desalination is an important source of potable water,
  necessary in order to keep up with demand generated by a rapidly growing global
  population. Although RO is currently the cheapest form of desalination, the
  cost must be further decreased in order to meet demand. By designing better
  membranes for desalination, we can achieve higher water fluxes and lower feed
  pressure requirements which will decrease the cost of fresh water production.
  \cite{humplik_nanostructured_2011}.

@@

  NF was introduced as an intermediate between RO and ultrafiltration, having
  the ability to separate organic matter and salts on the order of one nanometer
  in size. Explicit pore pathways running through the thickness of the membrane
  permit high solvent fluxes at lower pressures than RO. NF membranes typically
  have a surface charge resulting from ionizable groups which affords separation
  of ions smaller than the pore readius by Donnan exclusion.
  \cite{van_der_bruggen_review_2003} NF is often used as a precursor to reverse
  osmosis since it can efficiently remove a significant portion of dissolved
  solutes. Following NF pretreatment, RO can further purify the permeate. 
  Unfortunately, NF membranes, like RO, possess a pore size
  distribution which limits their ability to perform precise separations
  \cite{bowen_modelling_2002}. An NF membrane with high selectivity could offer
  the versatility to separate ions as well as small charged and uncharged
  molecules to the same degree as RO and at a significantly lower cost.

@@

  Development of nanostructured materials has been limited by the ability to
  synthesize and scale various fundamentally sound technologies.  Graphene sheets
  are atomically thick which results in excellent water permeability but defects
  during manufacturing severely impact selectivity.
  \cite{cohen-tanugi_multilayer_2016} Molecular dynamics simulations of carbon
  nanotubes show promise \cite{humplik_nanostructured_2011} but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity.
  \cite{hata_water-assisted_2004,maruyama_growth_2005} Zeolites have sub-nm pores
  with a narrow pore size distribution and MD simulations exhibit complete
  rejection of solvated ions, \cite{murad_molecular_1998} however, experimental
  rejection was low and attributed to interstitial defects formed during membrane
  synthesis. \cite{li_desalination_2004} There is a need for a scalable
  nanostructured membrane.

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer named
  Na-GA3C11 have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase \ref{fig:assembly} The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemsitry.

  Research into LLC membranes has been revived in recent years.  During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al. showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place. \cite{feng_scalable_2014} In 2016,
  Feng et al. showed that the same result could be obtained using a technique
  termed soft confinement. \cite{feng_thin_2016}

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. Rejection studies show that
  membranes formed by NA-GA3C11 can not perform separations of solutes less than
  1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
  We do not yet understand how to controllably reduce the effective pore size or
  how to tune the chemical environment in the nanopores for effective water
  desalination or small organic molecule separations. The only source of
  predictive modeling for LLC systems have been macroscopic models which likely
  do not adequately describe transport at these length scales.
  \cite{hatakeyama_water_2011} It will be challenging to efficiently narrow down
  the large design space in a laboratory setting without a robust model.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must gain a thorough
  understanding of the nanoscopic structure of LLC membranes.  Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental Wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate which will be laid out and numbered in subsequent paragraphs.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer. Their estimation is based on a simulated
  system containing only 16 total monomers which likely does not sufficiently
  model the chemical environment present in the real system.
  \cite{zhu_methacrylated_2006}.  A separate calculation based on the volume of
  the liquid crystal monomers proposes that there are seven monomers in each
  layer \cite{resel_structural_2000}.  A molecular model orders of magnitude
  larger than any other reported atomistic liquid crystal membrane simulations
  has the best chance of directly answering this question.  We will directly
  change the layer composition and note its effect on membrane structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system. \cite{gazit_possible_2002} Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. We can use simulated X-ray
  diffration patterns to compare the two stacking configurations.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically will
  require many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally and why others are not.

@@

  While the Col\textsubscript{h} phase is described as dry, it is has been
  suggested by experimentalists in unpublished communications that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a "dry" and "wet" system.

@@

@@

  %BJC: Would it be better to describe each reflection as a numbered list (like with the questions above?)
  We used experimental wide angle X-ray scattering (WAXS) data (produced as 
  described in \cite{feng_scalable_2014}) and small-angle X-ray scattering (SAXS)
  data from \cite{feng_thin_2016} to inform our choices of some initial
  structural parameters (Figure~\ref{fig:SWAXS}). We rely primarily on the 2D
  WAXS data since it encodes all structural details down to the sub-nm scale.
  There are five major features of interest present in the 2D experimental
  pattern shown in Figure \ref{fig:WAXS}. The first is located at $q_z$ = 1.7
  \angstrom$^{-1}$, corresponding to a real space separation of 3.7 \angstrom~.
  The reflection is attributed to $\pi$-$\pi$ stacking between aromatic rings in
  the direction perpendicular to the membrane plane, or z-axis
  \cite{feng_scalable_2014}. For simplicity, this reflection will be referred to
  as R-$\pi$. A weak intensity line is located at exactly half the $q_z$ value of
  R-$\pi$ ($q_z$ = 0.85 \angstrom$^-1$), corresponding to a real space periodic
  spacing of 7.4 \angstrom~. This reflection has been interpreted as
  2\textsubscript{1} helical ordering of aromatic rings along the z axis meaning
  if the positions of the aromatic rings can be traced by a helix, then for each
  turn in the helix, there should be two aromatic rings. For this reason it will
  be referred to as R-helix. A third major reflection is marked by a low
  intensity ring located at r = 1.4 \angstrom$^-1$. The real space separation
  corresponds to 4.5 \angstrom~ which is characteristic of the average spacing
  between packed alkane chains. This reflection will be called R-alkanes. Within
  R-alkanes, are four spots of higher relative intensity which will be called
  R-spots. All are located $\approx 40$ degrees from the $q_z$ axis in their
  respective quadrants. In many liquid crystal systems this can be explained by
  the tilt angle of the alkane chains with respect to the xy plane.  The final
  feature corresponds to the spacing and symmetry of the d\textsubscript{100}
  plane which can be related to the distance between pores.  The feature, which
  will be called R-pores, is characterized by dots along $q_z$ = 0. The spacing
  between dots is indicative of the hexagonal symmetry of the packed pores. The
  same information at higher resolution is obtained using a SAXS setup. By
  radially integrating the 2D data one gets a 1D curve which is shown in
  Figure~\ref{fig:SAXS}.

  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
        %       \vspace{12mm}
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \raisebox{.2\textwidth}{%
                \includegraphics[width=\linewidth]{WAXS_annotated.png}
                }
                \caption{}\label{fig:WAXS}
        \end{subfigure}
        \caption{(a) 1D small angle X-ray scattering indicates hexagonal packing of
        pores as well as the spacing between pores (b) 2D wide angle X-ray scattering
        gives details about repeating features less than 1 nanometer apart}\label{fig:SWAXS}
  \end{figure}

  \section*{Methods}

  \subsection*{Monomer Placement}

@@

  \subsection*{Unit Cell Preparation}

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic
  simulation, calling for a more efficient way to build the system.  Previous
  work has shown a coarse grain model self assemble into the H\textsubscript{II}
  phase configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (See supplemental information
  for further details). To bypass the slow self-assembly process, python scripts
  are used to assemble monomers into a structure close to one of a number of
  hypothesized equilibrium configurations (Figure~\ref{fig:python}).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. This point will be explained in more detail later.

\subsection*{Monomer Placement}

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm. The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will likely be unintended high energy repulsions during early
  equilibration. We chose an initial pore spacing of 4.5 nm, $\approx$ 10 \%
  larger than the experimental value of 4.12 nm.  The distance between layers,
  the rotation of the layers with respect to adjacent layers, and the number of
  monomers per layer do influence the equilibrium structure and require further
  justification for their choices. We rely on experimental data to inform our
  choices

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction.  A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. There will be no further discussion of this system but the interested
  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also unfeasible for the monomers to orient in the t-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  As outlined in \ref{point:monomernum} the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

  \subsection*{Equilibration}

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 1e6 KJ mol$^-1$ nm$^-2$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^-1$ nm$^-2$, the restraints are released linearly until there is no more
  restraining potential. The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run a long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time consant of 10 ps.
  In order to create a wet system, we solvated an initial configuration with
  water using gmx solvate. All water molecules placed outside the pore region are
  removed. Waters inside the pore region are randomly removed until the desired
  concentration of water in the pores is reached. The remainder of the
  equilibration follows the same procedure as the dry system. In all cases, the
  v-rescale thermostat was used with tau-t = 0.1. % Supplemental

>>>>

  \section{Introduction}

  Nanostructured membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control membrane architecture at the atomic scale
  permitting the design of solute-specific separation membranes.
  \cite{humplik_nanostructured_2011} Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

  Current commercial RO membranes are thin film composite membranes with a
  porous support layer and an active layer made of a dense polymer matrix. The
  membranes are unstructured with tortuous and polydisperse diffusion pathways.
  Separation occurs based on differences in solubility and diffusivity of solutes
  within the polymer matrix \cite{wijmans_solution-diffusion_1995}. With
  optimization, one can exploit these differences to create a functional
  selective barrier. RO operation costs are suboptimal because high feed
  pressures are required in order to achieve a useful solvent flux.

  %BJC : I don't know that this paragraph fits.  
  Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
  the world's water is fresh. Desalination is an important source of potable
  water, necessary in order to keep up with demand generated by a rapidly growing
  global population. Although RO is currently the cheapest form of desalination,
  the cost must be further decreased in order to meet demand. A typical seawater
  desalination process requires feed pressures in the range of 800 to 1000 psi
  \cite{fritzmann_state---art_2007}. By designing better membranes for
  desalination, we can achieve higher water fluxes and reduce feed pressure
  requirements which will decrease the cost of fresh water production.
  \cite{humplik_nanostructured_2011}.

@@

  NF was introduced as an intermediate between RO and ultrafiltration, having
  the ability to separate organic matter and salts on the order of one nanometer
  in size. Explicit pore pathways running through the thickness of the membrane
  permit high solvent fluxes at lower pressures than RO. NF membranes typically
  have a surface charge resulting from ionizable groups which affords separation
  of ions smaller than the pore radius by the mechanism of Donnan exclusion
  \cite{van_der_bruggen_review_2003}. NF is often used as a precursor to reverse
  osmosis since it can efficiently remove a significant portion of dissolved
  solutes. Following NF pretreatment, RO can further purify the permeate. 
  Unfortunately, NF membranes, like RO, possess a pore size
  distribution which limits their ability to perform precise separations
  \cite{bowen_modelling_2002}. An NF membrane with high selectivity could offer
  the versatility to separate ions as well as small charged and uncharged
  molecules to the same degree as RO and at a significantly lower cost.

@@

  Development of nanostructured materials has been limited by the ability to
  synthesize and scale various fundamentally sound technologies. Graphene sheets
  are atomically thick which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}. Molecular dynamics (MD) simulations of carbon
  nanotubes show promise \cite{humplik_nanostructured_2011} but synthetic
  techniques are unable to achieve scalable alignment and pore monodispersity.
  \cite{hata_water-assisted_2004,maruyama_growth_2005} Zeolites have sub-nm pores
  with a narrow pore size distribution and MD simulations exhibit complete
  rejection of solvated ions, \cite{murad_molecular_1998} however, experimental
  rejection was low and attributed to interstitial defects formed during membrane
  synthesis. \cite{li_desalination_2004} There is a need for a scalable
  nanostructured membrane.

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemsitry.

  Research into LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al. showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 can not perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

@@

  %BJC: I think the first sentence of the following paragraph is
  %necessary to put somewhere, but maybe not in this spot exactly
  In order to appropriately model transport, we must first gain a thorough
  understanding of the nanoscopic structure of LLC membranes. Our approach to
  constructing a general model will follow the development of a model of the
  assembly formed by Na-GA3C11 since it has sufficient experimental
  characterization. We have also narrowed our scope to the development of a model
  of the Col\textsubscript{h} phase membrane. Compared to the H\textsubscript{II}
  phase, the Col\textsubscript{h} phase is a simpler starting point, due to the
  absence of water, and has detailed experimental wide-angle X-ray scattering
  (WAXS) patterns useful for reconstructing structural data.

  Despite having structural data, there is still information which experiment
  cannot definitively answer. There are several key questions that we will
  investigate.
% which will be laid out and numbered in subsequent paragraphs.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer. Their estimation is based on a simulated
  system containing only 16 total monomers which likely does not sufficiently
  model the chemical environment present in the real system
  \cite{zhu_methacrylated_2006}.  A separate calculation based on the volume of
  the liquid crystal monomers proposes that there are seven monomers in each
  layer \cite{resel_structural_2000}.  A molecular model orders of magnitude
  larger than any other reported atomistic liquid crystal membrane simulations
  has the best chance of directly answering this question.  We will directly
  change the layer composition and note its effect on membrane structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. We will compare simulated X-ray
  diffration patterns to experiment in order to deduce which stacking 
  configurations is most likely.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically will
  require many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a "dry" and "wet" system.

@@

  %BJC: Would it be better to describe each reflection as a numbered list (like with the questions above?)
  We used experimental small-angle X-ray scattering (SAXS) data from
  \cite{feng_thin_2016} (Fig. ~\ref{fig:SAXS}) and wide angle X-ray scattering
  (WAXS) data (Fig. ~\ref{fig:WAXS}, produced as described in
  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
  since it encodes all structural details down to the sub-nm scale.  There are
  five major features of interest present in the 2D experimental pattern shown in
  Figure \ref{fig:WAXS}. The first is located at $q_z$ = 1.7 \angstrom$^{-1}$,
  corresponding to a real space separation of 3.7 \angstrom~.  The reflection is
  attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
  perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For
  simplicity, this reflection will be referred to as R-$\pi$. A weak intensity
  line is located at exactly half the $q_z$ value of R-$\pi$ ($q_z$ = 0.85
  \angstrom$^{-1}$), corresponding to a real space periodic spacing of 7.4
  \angstrom~. This reflection has been interpreted as 2\textsubscript{1} helical
  ordering of aromatic rings along the z axis meaning if the positions of the
  aromatic rings can be traced by a helix, then for each full turn in the helix, one
  will encounter two aromatic rings. For this reason it will be referred to as
  R-helix. A third major reflection is marked by a low intensity ring located at
  r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to 4.5 \angstrom~
  which is characteristic of the average spacing between packed alkane chains.
  This reflection will be called R-alkanes. Within R-alkanes, are four spots of
  higher relative intensity which will be called R-spots. All are located
  $\approx 37$ degrees from the $q_z$ axis in their respective quadrants. In many
  liquid crystal systems this can be explained by the tilt angle of the alkane
  chains with respect to the membrane plane. The final feature corresponds to the
  spacing and symmetry of the d\textsubscript{100} plane which can be related to
  the distance between pores.  The feature, which will be called R-pores, is
  characterized by dots along $q_z$ = 0. The spacing between dots is indicative
  of the hexagonal symmetry of the packed pores. The same information at higher
  resolution is obtained using a SAXS setup.

@@

  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

  \section{Methods}

  \subsection{Monomer Parameterization}

@@

  \subsection{Unit Cell Preparation}

  The timescale for self assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system.  Previous work has shown
  a coarse grain model self assemble into the H\textsubscript{II} phase
  configuration in $\approx$ 1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system To bypass the slow
  self-assembly process, python scripts are used to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.

\subsection{Monomer Placement}

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental). The pore radius is chosen to be
  0.6 nm in our initial configurations because the pore size is estimated to be
  $\approx$ 1.2 nm. The initial distance between pores also has little effect on
  the the equilibrated structure. However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\approx$ 10 \% larger than the experimental
  value of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. There will be no further discussion of this system but the interested
  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also unfeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  As outlined in (\ref{point:monomernum}) the number of monomers in each layer is
  unknown. We tested configurations constructed with a varied number of
  monomers per layer. Systems were built in the offset and parallel displaced
  configurations with 4, 5, 6, 7 and 8 monomers per layer.

  \subsection{Equilibration}

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 1e6 KJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are released linearly until there is no more
  restraining potential. The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time consant of 10 ps.

  In order to create a wet system, we solvated an initial configuration with
  water using gmx solvate. All water molecules placed outside the pore region are
  removed. Waters inside the pore region are randomly removed until the desired
  concentration of water in the pores is reached. The remainder of the
  equilibration follows the same procedure as the dry system.

  \subsection{Crosslinking}

  % BJC: Need to make it clear that this was not done on all systems
